### 2023-10-09 scripts/analysis.Rmd

---
title: "PREVIC Item selection"
output: html_document
date: "2023-07-20"
---

# Packages

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(ggthemes)
library(tidybayes)
library(brms)
library(rstan)
library(loo)
library(coda)
library(testing)
library(ggridges)
library(lavaan)
library(ggbreak)
library(grid)
library(cowplot)

estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}

func <- function(x){
  abs(1-x)
}
```

# Data

```{r}
data <- read_csv("../data/previc_data.csv")%>%
  mutate(age_group = factor(substr(age, 1,1)))

irt_dat <- data%>%
  select(subjID, word, score, sex, aoa_rating_german)

aoa <- data%>%distinct(word, .keep_all = T)%>%select(word, aoa_rating_german)

full <- data%>%
  group_by(subjID)%>%
  summarise(mean_full = mean(score))

```

# Item selection

Models take a long time to run and generate files that are too large to be saved on github. In the script below, all code that takes long to run or depends on large files that need to be run beforehand is commented out. Only aggregated that are part of the repository are not commented out.

## Step 1: Selection based on In- and Outfit

### 1PL model with all items

```{r}
prior_1pl <- 
  prior("normal(0, 1)", class = "sd", group = "subjID") + 
  prior("normal(0, 3)", class = "sd", group = "word")
```

#### Model

```{r}
# irt1 <- brm(
#   data = irt_dat,
#   family = bernoulli(),
#   score ~ 1 + (1 | word) + (1 | subjID),
#   prior = prior_1pl,
#   control = list(adapt_delta = 0.95, max_treedepth = 20),
#   cores = 6,
#   chains = 6,
#   iter = 6000,
#   threads = threading(8), #to speed things up, comment out if not on a cluster
#   backend = "cmdstanr" #to speed things up, comment out if not on a cluster
# )
# 
# saveRDS(irt1, "../saves/irt1.rds")
# 
# irt1 <- add_criterion(
#   irt1,
#   criterion = c("loo","waic"),
#   cores = 1,
#   ndraws = 2000
# )
# 
# saveRDS(irt1, "../saves/irt1.rds")
# 
# irt1 <- readRDS("../saves/irt1.rds")
```

#### Extract In- and Outfit

```{r}
# rasch_fit_draws <- irt_dat%>%
#   add_epred_draws(irt1, re_formula = ~(1 | word) + (1 | subjID), ndraws = 1000)
#   
# rasch_fit <- rasch_fit_draws%>%
#   mutate(zvi = (score - .epred)/(.epred*(1-.epred))^0.5)%>%
#   group_by(word,aoa_rating_german, .draw)%>%
#   summarise(outfit = sum(zvi^2)/length(unique(subjID)),
#             infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))
# 
# rasch_fit_mode <- rasch_fit%>%
#   pivot_longer(names_to = "fit_index", values_to = "value", cols = c(outfit, infit))%>%
#   group_by(word, aoa_rating_german, fit_index)%>%
#   summarise(mode = estimate_mode(value),
#             lci = hdi_lower(value),
#             uci = hdi_upper(value))
# 
# saveRDS(rasch_fit_mode, "../saves/rasch_fit_mode.rds")

rasch_fit_mode <- readRDS("../saves/rasch_fit_mode.rds")

```

#### Select items

```{r}
# remove items that have very poor in and outfit

# fit_selected_items <- rasch_fit_mode%>%ungroup()%>%filter(0.7 < mode & 1.3 > mode)%>%group_by(word)%>%mutate(n = n())%>%filter(n == 2)%>%distinct(word)%>%pull(word)
# 
# saveRDS(fit_selected_items, "../saves/fit_selected_items.rds")

fit_selected_items <- readRDS( "../saves/fit_selected_items.rds")

```

## Step 2: Compute indices for items selected in step 1

### 1PL model for selected items

Run script `run1PLmodel_sel_fit.R`

```{r}
# irt1_fit_sel <- readRDS("../saves/irt1_fit_sel.rds")
#
# ranef(irt1_fit_sel)$word%>%as_tibble(rownames = "word")%>%saveRDS("../saves/easiness_1PL_sel.rds")

```

#### Extract In- and Outfit

```{r}
# rasch_fit_draws_fit_sel <- irt_dat%>%
#   filter(word %in% fit_selected_items)%>%
#   add_epred_draws(irt1_fit_sel, re_formula = ~(1 | word) + (1 | subjID), ndraws = 1000)
# 
# rasch_fit_fit_sel <- rasch_fit_draws_fit_sel%>%
#   mutate(zvi = (score - .epred)/(.epred*(1-.epred))^0.5)%>%
#   group_by(word,aoa_rating_german, .draw)%>%
#   summarise(outfit = sum(zvi^2)/length(unique(subjID)),
#             infit = (sum(zvi^2*(.epred*(1-.epred)))/sum(.epred*(1-.epred))))
# 
# rasch_fit_mode_fit_sel <- rasch_fit_fit_sel%>%
#   pivot_longer(names_to = "fit_index", values_to = "value", cols = c(outfit, infit))%>%
#   group_by(word, aoa_rating_german, fit_index)%>%
#   summarise(mode = estimate_mode(value),
#             lci = hdi_lower(value),
#             uci = hdi_upper(value))
# 
# saveRDS(rasch_fit_mode_fit_sel, "../saves/rasch_fit_mode_fit_sel.rds")

rasch_fit_mode_fit_sel <- readRDS("../saves/rasch_fit_mode_fit_sel.rds")
```

### Compute modindices based on frequentist model

```{r}
# all_items <- irt_dat%>%
#   filter(word %in% fit_selected_items)%>%
#   distinct(word)%>%
#   pull(word)
# 
# irt_all <- irt_dat%>%
#   filter(word %in% fit_selected_items)%>%
#   select(-aoa_rating_german)%>%
#   pivot_wider(names_from = word, values_from = score)%>%
#   select( -subjID, -sex)
# 
# modelAllx <- paste(paste0("1*",all_items, "+"), collapse = " ")
# 
# modelAll <- paste0("f =~", substr(modelAllx, 1, nchar(modelAllx)-1))
# 
# freqAll <- sem(modelAll, irt_all, ordered =TRUE, parameterization = "theta")
# 
# saveRDS(freqAll, "../saves/freqAll.rds")
# 
# freqAll <- readRDS("../saves/freqAll.rds")
# 
# miAll <- modindices(freqAll)
# 
# miAllSel <- miAll%>%
#   filter(lhs == "f")
#   
# saveRDS(miAllSel, "../saves/miAllSel.rds")
```

## Indices for item selection

```{r}
mi <- readRDS("../saves/miAllSel.rds")%>%arrange(rhs)%>%pull(mi)
items <-data%>%filter(word %in% readRDS( "../saves/fit_selected_items.rds"))%>%distinct(word)%>%arrange(word)%>%pull(word)
easiness_1PL_fit_sel <- readRDS("../saves/easiness_1PL_sel.rds")%>%arrange(word) %>%pull(Estimate.Intercept)
infit_fit_sel <- readRDS("../saves/rasch_fit_mode_fit_sel.rds") %>% filter(fit_index == "infit")%>%arrange(word)%>% pull(mode)
outfit_fit_sel <- readRDS("../saves/rasch_fit_mode_fit_sel.rds") %>% filter(fit_index == "outfit")%>%arrange(word)%>% pull(mode)
```

## Step 3: Simulated annealing

```{r}
source("../scripts/simulated_annealing.R")
```

### Test run

```{r, message=F, warning=F, comment=F}
sim_test <- simulated_annealing_rasch(100)

items[unlist(sim_test$best_subset) == TRUE]
```

### Compare 1PL and 2PL model for different sizes

Run script `run2PLmodel_sel_fit.R` to fit 2PL model with selected items (needed for model comparison)

```{r}
irt2PL_fit_sel <- readRDS("../saves/irt2PL_fit_sel.rds")
```

Run script `selected_model_comparison.R`

#### Visualize Model comparison for different sizes

```{r}
readRDS("../saves/model_comparison_size.rds")%>%
  mutate(ratio = abs(elpd_diff) / (2*se_diff))%>%
  mutate(ratio = ifelse( elpd_diff == 0,0,ratio))%>%
  filter(ratio != 0)%>%
  mutate(ratio = ifelse(model == "m2PL", ratio*-1, ratio))%>%
  arrange(iter, size)%>%
  ggplot(aes(y = size, x = ratio, col = factor(iter)))+
  geom_vline(xintercept = c(-1,1), alpha = .5, col = "#31493C", lty = 3)+
  geom_vline(xintercept = 0, col = "black", lty = 1, size = 1)+
  geom_point(alpha = .75)+
  scale_y_continuous(breaks = c(70,75,80,85,90,95,100,125,175))+
  scale_x_continuous( limits = c(-6,6), breaks = c(-5,-2, -1, 0, 1, 2, 5), labels = c("5.00","2.00","1.00","0.00", "1.00","2.00", "5.00"))+
  scale_color_ptol(guide = "none")+
  theme_bw()+
  theme(panel.grid.minor = element_blank())+
  labs(y = "No. of items in subset", x = expression(paste("Model comparison: ", frac(Delta ~ elpd, "2 *"~SE(Delta ~ elpd)))))+
  annotation_custom(textGrob("1PL wins", 
                             gp=gpar(fontsize=13,
                                     col = "black", 
                                     fontface="bold")),
                    xmin=-3, xmax=-3, ymin=160, ymax=160) +
  annotation_custom(textGrob("2PL wins",
                             gp=gpar(fontsize=13,
                                     col = "black",
                                     fontface="bold")),
                    xmin=3, xmax=3, ymin=160, ymax=160)

```

### Select items for preferred size (90 items)

Run script `item_selection.R`

```{r}
sel_items_90 <- readRDS("../saves/item_selection.rds")%>%
  filter(size == 90)%>%
  select(items)%>%
  unnest()%>%
  group_by(items)%>%
  summarise(n = n())%>%
  arrange(-n)%>%
  head(90)%>%
  pull(items)

```

## Step 4: Differential item functioning (split by sex)

### Model

```{r}
irt_dat_sel_90 <- irt_dat%>%
  filter(word %in% sel_items_90)
```

```{r}
# irt1_final_90_dif_sex <-  brm(
#   data = irt_dat_sel_90,
#   family = bernoulli(),
#   score ~ 1 + (0+ sex | word) + (1 | subjID),
#   prior = prior_1pl,
#   control = list(adapt_delta = 0.95, max_treedepth = 20),
#   cores = 6,
#   chains = 6,
#   iter = 8000,
#   threads = threading(8), #to speed things up, comment out if not on a cluster
#   backend = "cmdstanr" #to speed things up, comment out if not on a cluster
# )%>%add_criterion(c("loo"), cores = 2, ndraws = 2000)
# 
# 
# saveRDS(irt1_final_90_dif_sex, "../saves/irt1_final_90_dif_sex.rds")
# 
# irt1_final_90_dif_sex<- readRDS("../saves/irt1_final_90_dif_sex.rds")

```

### Visual model inspection

```{r}
# final_90_dif_sex <- as_draws_df(irt1_final_90_dif_sex)%>%
#   select(b_Intercept, starts_with("r_word"))%>%
#   mutate(iter = 1:n()) %>%
#   pivot_longer(starts_with("r_word")) %>%
#   mutate(name = str_remove(name, pattern = "r_word"),
#          name = str_remove_all(name, pattern = "\\[|\\]"))%>%
#   separate(name, into = c("item", "sex"), sep = "\\,")%>%
#   mutate(sex = str_remove(sex, pattern = "sex"),
#          val =  value)%>%
#   group_by(item, sex)%>%
#   summarise(mode = estimate_mode(val),
#             uci = hdi_upper(val),
#             lci = hdi_lower(val))%>%
#   left_join(aoa%>%rename(item = word))
# 
# saveRDS(final_90_dif_sex, "../saves/model_params_irt1_final_90_dif_sex.rds")

final_90_dif_sex <- readRDS("../saves/model_params_irt1_final_90_dif_sex.rds")

final_90_dif_sex%>%
  group_by(item)%>%
  mutate(dif = abs(abs(mode) - lag(abs(mode))))%>%
  fill(dif, .direction = "up")%>%
ggplot(.,aes(x = reorder(item, dif))) +
	geom_point(aes(col = sex, y = lci), position = position_dodge(width = .5)) +
  geom_point(aes(col = sex, y = uci), position = position_dodge(width = .5)) +
  geom_linerange(aes(col = sex, ymin = lci + 0.1, ymax = uci-0.1), position = position_dodge(width = .5), alpha = .5) +
	coord_flip() +
  scale_color_colorblind(labels = c("male","female"), name = "Group")+
	labs(x = "Item", y = "Easiness estimate")+
  theme_minimal()
```
```{r}
final_90_dif_sex%>%
  pivot_wider(names_from = sex, values_from = c(mode,uci,lci))%>%
  ggplot(., aes(x = mode_f, y = mode_m))+
  geom_abline(intercept = 0, slope = 1, lty = 3, alpha = .75)+
  geom_point(pch = 1, size = 2, stroke  = 1, aes(col = factor(aoa_rating_german)))+
  geom_linerange(aes(ymin = lci_m, ymax = uci_m),  alpha = .25, lty = 1)+
  geom_linerange(aes(xmin = lci_f, xmax = uci_f),  alpha = .25, lty = 1)+
  #geom_text(aes(label = item, x = uci_f +0.2))+
  labs(x = "Group: female", y = "Group: male")+
  scale_color_viridis_d()+
  guides(col = F)+
  coord_fixed()+
  theme_few()

```

## Final item selection

```{r}
# final_items <- readRDS("../saves/item_selection.rds")%>%
#   filter(size == 90)%>%
#   select(items)%>%
#   unnest()%>%
#   group_by(items)%>%
#   summarise(n = n())%>%
#   arrange(-n)%>%
#   head(90)%>%
#   filter(items != "verloben")%>%
#   pull(items)
# 
# saveRDS(final_items, "../saves/final_items.rds")
```

```{r}
data%>%
  filter(word %in% readRDS("../saves/final_items.rds"))%>%
  distinct(word, english, word_type)%>%
  write_csv("../data/final_item_list.csv")
```

# Final model

## Model

```{r}
irt_dat_final <- irt_dat%>%
  filter(word %in% final_items)
```

```{r}
# irt1_final <- brm(
#   data = irt_dat_final,
#   family = bernoulli(),
#   score ~ 1 + (1| word) + (1 | subjID),
#   prior = prior_1pl,
#   control = list(adapt_delta = 0.95, max_treedepth = 20),
#   cores = 6,
#   chains = 6,
#   iter = 8000,
#   threads = threading(8), #to speed things up, comment out if not on a cluster
#   backend = "cmdstanr" #to speed things up, comment out if not on a cluster
# )%>%add_criterion(c("loo"), cores = 2, ndraws = 2000)
# 
# 
# saveRDS(irt1_final, "../saves/irt1_final.rds")
# 
# irt1_final <- readRDS("../saves/irt1_final.rds")
```

## Item characteristics curves

```{r}
# icc1_final <- posterior_samples(irt1_final)%>% 
#   select(b_Intercept, starts_with("r_word"))%>%
#   mutate(iter = 1:n()) %>% 
#   pivot_longer(starts_with("r_word"), names_to = "item", values_to = "xi") %>%
#   mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
#   expand(nesting(iter, b_Intercept, item, xi),
#          theta = seq(from = -6, to = 6, length.out = 100)) %>% 
#   mutate(p = inv_logit_scaled(b_Intercept + xi + theta)) %>% 
#   group_by(theta, item) %>% 
#   summarise(p = mean(p))%>%
#   left_join(aoa%>%rename(item = word))
# 
# saveRDS(icc1_final, "../saves/icc1_final.rds")

icc1_final <- readRDS("../saves/icc1_final.rds")
```

```{r}
p_icc <- icc1_final %>% 
  ggplot(aes(x = theta, y = p,group = item, col = aoa_rating_german)) +
  geom_line() +
  scale_color_viridis_c(name = "Rated age of acquisition") +
  labs(x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  theme_few()+
  theme(legend.position = c(0.85, 0.2), legend.direction = "horizontal", legend.title.align = 0.5, legend.background = element_blank())+
  guides(colour = guide_colourbar(title.position="top"))
```

## Test information curve

```{r}
# tic1_final <- as_draws_df(irt1_final) %>% 
#   select(.draw, b_Intercept, starts_with("r_word")) %>% 
#   pivot_longer(starts_with("r_word"), names_to = "item", values_to = "xi") %>%
#   mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
#   expand(nesting(.draw, b_Intercept, item, xi),
#          theta = seq(from = -6, to = 6, length.out = 200)) %>% 
#   mutate(p = inv_logit_scaled(b_Intercept + xi + theta)) %>% 
#   mutate(i = p * (1 - p)) %>% 
#   group_by(theta, .draw) %>% 
#   summarise(sum_i = sum(i)) %>% 
#   group_by(theta) %>% 
#   summarise(i = median(sum_i))
#   
# saveRDS(tic1_final, "../saves/tic1_final.rds")

tic1_final <- readRDS("../saves/tic1_final.rds")

p_tic <- ggplot(tic1_final, aes(x = theta, y = i)) +
  geom_line() +
  labs(x = expression(theta~('ability on the logit scale')),
       y = "Test information") +
  theme_few()
```

```{r}
ggdraw() +
  draw_plot(p_icc)+
  draw_plot(p_tic + theme(plot.background = element_blank()), x = 0.065, y = .68, width = .25, height = .3)

```
## Reliability

### KR20

```{r}
rel_dat <- irt_dat%>%
  select(-sex, -aoa_rating_german)%>%
  filter(word %in% final_items)%>%
  group_by(subjID)%>%
  distinct(word, .keep_all = T)%>%
  pivot_wider(names_from = word, values_from = score)%>%
  ungroup()%>%
  select(-subjID)

kr20_rel <- kr20(rel_dat, hit = 1)

kr20_rel
```

### Andrich Reliability

```{r}
# pers_params <- ranef(irt1_final)$subjID%>%as_tibble(rownames = "subjID")
# 
# andrich_rel <- 1 - 
#   (1/length(pers_params$Estimate.Intercept) * sum(pers_params$Est.Error.Intercept^2))/
#   (1/(length(pers_params$Estimate.Intercept)-1)*sum((pers_params$Estimate.Intercept - mean(pers_params$Estimate.Intercept))^2))
# 
# andrich_rel
```

```{r}
# tibble(type = c("kr20", "andrich"), 
#        rel = c(kr20_rel, andrich_rel))%>%
#   saveRDS("../saves/reliability.rds")

readRDS("../saves/reliability.rds")
```

# Validity

Link PREVIC scores with receptive vocabualry scores of the oREV.

```{r}
orev_data <- read_csv("../data/orev_data.csv")

link_data <- data%>%
  filter(word %in% final_items)%>%
  group_by(subjID)%>%
  summarise(previc = sum(score))%>%
  left_join(orev_data%>%
              group_by(subjID)%>%
              summarise(orev = sum(correct)))%>%
  filter(!is.na(orev))%>%
  ungroup()%>%
  distinct(subjID, .keep_all = T)

cor.test(link_data$previc, link_data$orev)
```

```{r}
ggplot(link_data, aes(x = orev, y = previc))+
  geom_smooth(method = "lm", inherit.aes = F, aes(y = previc, x = orev), col = "firebrick")+
  geom_point(pch = 19,stroke = F,  alpha = .5, size = 2)+
  stat_cor(inherit.aes = F, aes(x = previc, y = orev))+
  theme_bw()
```

```{r}
m_prev_orev <- data%>%
  filter(word %in% final_items)%>%
  group_by(subjID)%>%
  group_by(age, subjID)%>%
  summarise(sum  = sum(score),
            n = n())%>%
  left_join(orev_data%>%
              group_by(subjID)%>%
              summarise(orev = sum(correct)))%>%
  filter(!is.na(orev))%>%
  ungroup()%>%
  mutate(age = scale(age),
         orev = scale(orev))

m1 <- brm(sum |trials(n) ~ age + orev, family = binomial, data = m_prev_orev, chains = 4, cores = 4)

m1
```